\documentclass[twoside]{article}
\usepackage{mathrsfs}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{dcolumn}
\usepackage{subfigure}
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{ulem}

\newcommand{\myheader}[1]{\textcolor{black}{\textbf{#1}}}
\setlength{\parindent}{0in}

\begin{document}
	
	{\bf Advanced Topics in Development \sout{and International Trade}} \hfill {\bf
		EEP 152 : Section Syllabus}
	
	\vspace{0.5em}
	
	\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}ll@{\extracolsep{\fill}}l}
		\hline
		\multicolumn{4}{c}{} \\ [-0.5em]
		{\bf Professor} & Jeremy Magruder & {\bf GSI} & John Loeser \\
		{\bf GSI OH} & W 10-11, F 10-11, or by appt. & {\bf GSI OH location} & Giannini 234 \\
		{\bf Section 1 time} & W 3-4 & {\bf Section 1 location} & Morgan 138 \\
		{\bf Section 2 time} & F 9-10 & {\bf Section 2 location} & Giannini 201 \\
		& & {\bf email} & \href{mailto:john.a.loeser@berkeley.edu}{john.a.loeser@berkeley.edu} \\
		\multicolumn{4}{c}{} \\ [-0.5em]
		\hline
	\end{tabular*}
	\vspace{10pt} \bigskip
	
	The purpose of section (once weekly, 1 hour) is to reinforce the material from the course. I'll spend the first half of each section reviewing material from lecture or techniques for homework and exams, and we'll spend the second half in guided discussion of the course material. \\
	
	The section syllabus and notes will all be available at \href{http://www.github.com/johnloeser/eep152}{www.github.com/johnloeser/eep152}.
	
	\vspace{10pt}
	\begin{tabular}{ll}
		\bf{Aug 31}  &  Introduction, Solow Model \\
		\bf{Sep 7, 14, 21, 28} &  TBA \\
		\bf{Oct 5} &  No section \\
		\bf{Oct 12, 19, 26, Nov 2, 9, 16} &  TBA \\
		\bf{Nov 23} &  No section (Thanksgiving) \\
		\bf{Nov 30, Dec 7} &  TBA \\
	\end{tabular}
	%\newpage
	\bigskip
	
	{\bf Attendance}: Try to attend the section you've signed up for. If you need to switch, let me know. Attendance isn't mandatory, but section should be useful so you should attend. If it isn't useful, let me know and we'll make it useful (see Feedback). \\
	
	{\bf E-mail policy}: Please begin the subject with [EEP 152]. I'll do my best to respond to all emails within a business day. If I don't respond within two business days, your email was probably buried, so sending me a quick reminder to reply would be helpful. \\
	
	{\bf Office hours}: Office hours (twice weekly, 1 hour) are for discussing anything related to course material, very loosely defined. It's a good opportunity to get more specific, personalized feedback. There are no OH W Aug 31, W Sep 7, and the weeks of Oct 5 (W Oct 5, F Oct 7) and Nov 23 (W Nov 23, F Nov 25). \\
	
	{\bf Feedback}: If you don't find section useful or would prefer something change, please let me know immediately and we can work to change it. You can leave anonymized feedback at \href{https://goo.gl/forms/IYlJK3FRreQRh6TV2}{https://goo.gl/forms/IYlJK3FRreQRh6TV2}. \\
	
	{\bf Special accommodations}: If you require any disability-related accommodations during our sections, lectures, or exams, please inform Jeremy as soon as possible. Generally, you have until the second week of class. For such accommodations, you must provide a supporting letter from the Disabled Students' Program (DSP), whose website is \href{http://dsp.berkeley.edu}{dsp.berkeley.edu}. \\
	
	{\bf Honor code}: As a member of the UC Berkeley community, I act with honesty, integrity, and respect for others.
	
\end{document}